---
title: "Hands-on Exercise 4"
author: "Fangyuan Zhu"
date: 4 Feb 2023
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

## 1 Get started

In this In-class exercise, we will explore more interesting packages to visualize data with statistical summary information shown in the plots.

## 2 Packages used

Packages will be used in this exercise are:

-   *plotly*

-   *tidyverse*

-   *DT*

-   *patchwork*

-   *ggstatsplot*

-   *PMCMRplus*

```{r}
install.packages("ggstatsplot")
pacman::p_load(plotly, tidyverse, DT, patchwork, ggstatsplot, PMCMRplus)
```

## 3 Data used

Data we will use in this exercise is *Exam_data.csv,* and we can import the data with codes chunk below.

```{r}
exam_data<-read_csv("data/Exam_data.csv")
```

## 4 Working with visual variable: *plot_ly()* method

Today we are going to explore the function `plot_ly()` to create an interactive plot.

```{r}
plot_ly(data = exam_data,
        x = ~ENGLISH,
        y = ~MATHS,
        color = ~RACE)
```

```{r}
p<-ggplot(data = exam_data,
          aes(x = MATHS,
              y = ENGLISH)) +
  geom_point(dotsize = 1) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100))

ggplotly(p)
```

## 5 Working with *ggstatsplot* package

```{r eval=FALSE}
ggbetweenstats(
  data = exam_data,
  x = GENDER,
  y = MATHS,
  type = "o",
  messages = FALSE
)
```
